---
title: "Homework 3"
author: "Marisa Sobel"
date: "10/7/2018"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(ggridges)
library(patchwork)
library(ggthemes)

theme_set(theme_bw() + theme(legend.position = "bottom"))
```

## Problem 1 

Data cleaning

```{r, collapse=TRUE}
# add library that includes dataset 
library("p8105.datasets")

# import brfss_smart2010 data, clean names, filter by "Overall Health", remove unwanted vars
# make "response" a factor variable

brfss_data = 
  brfss_smart2010 %>% 
  janitor::clean_names() %>% 
  filter(topic == "Overall Health") %>% 
  select(-class, -topic, -question, 
         -sample_size, -(confidence_limit_low:geo_location)) %>% 
  mutate(
    response = factor(response, 
                      levels = c("Excellent", "Very good", "Good", "Fair", "Poor"))
  )
```

### Questions

1. _*In 2002, which states were observed at 7 locations?*_

```{r, collapse=TRUE}
# Filter by year, group by state, summarize number of locations in each state

brfss_data %>% 
  filter(year == 2002) %>% 
  group_by(locationabbr) %>% 
  summarise(
    number = n_distinct(locationdesc)
  ) %>% 
  filter(number == 7)
```

**Connecticut, Florida, and North Caroline** were observed at 7 locations in 2002.

2. _*Make a “spaghetti plot” that shows the number of observations in each state from 2002 to 2010.*_

```{r, collapse=TRUE}
# Group by year and state, summarize number of locations in each state (n)
# Line plot of year vs n, turn off legend --> too many states

brfss_data %>% 
  group_by(year, locationabbr) %>% 
  summarise(
    number = n()) %>% 
  ggplot(aes(x = year, y = number, color = locationabbr)) +
  geom_line() +
   labs(
    title = "Number of Observations by State (2002-2010)", 
    x = "Year", 
    y = "Number of Obs", 
    legend = "State") +
  viridis::scale_color_viridis(discrete = TRUE) +
  theme(legend.position = "none")
```

3. _*Make a table showing, for the years 2002, 2006, and 2010, the mean and standard deviation of the proportion of “Excellent” responses across locations in NY State.*_

```{r, collapse=TRUE}
# Filter: 1) NY, 2) year (2002, 2006, 2010)
# Group by location 
# "data_value" = proportion 

brfss_data %>% 
  filter(locationabbr == "NY") %>% 
  filter(year == 2002 | year == 2006 | year == 2010) %>% 
  filter(response == "Excellent") %>% 
  group_by(year) %>% 
  summarise(
    mean_excellent = mean(data_value), 
    sd_excellent = sd(data_value)
  )
```

4. _*For each year and state, compute the average proportion in each response category (taking the average across locations in a state). Make a five-panel plot that shows, for each response category separately, the distribution of these state-level averages over time.*_

```{r, collapse=TRUE}
# Group by year, state, and response, mean of "data_value"
# Line plot year vs mean, facet by "response", turn off legend --> too many states 

brfss_data %>% 
  group_by(year, locationabbr, response) %>% 
  summarise(
    mean = mean(data_value, na.rm = TRUE)
  ) %>% 
  ggplot(aes(x = year, y = mean, color = locationabbr)) +
  geom_line(alpha = .5) +
  facet_grid(~ response) + 
  labs(
    title = "Average Percent of Each Response by State (2002-2010)", 
    x = "Year", 
    y = "Average Percent", 
    legend = "State") +
  viridis::scale_color_viridis(discrete = TRUE) +
  theme(legend.position = "none")
```


## Problem 2

```{r, collapse=TRUE}
# import dataset
instacart_data = instacart 
```

Data exploration 

```{r}

```

Short description, size and structure, key variables and illustrative examples of obs 

### Questions 

1. _*How many aisles are there, and which aisles are the most items ordered from?*_

```{r}

```

2. _*Make a plot that shows the number of items ordered in each aisle. Order aisles sensibly, and organize your plot so others can read it.*_

```{r}

```

3. _*Make a table showing the most popular item aisles “baking ingredients”, “dog food care”, and “packaged vegetables fruits”.*_

```{r}

```

4. _*Make a table showing the mean hour of the day at which Pink Lady Apples and Coffee Ice Cream are ordered on each day of the week; format this table for human readers (i.e. produce a 2 x 7 table).*_

```{r}

```


## Problem 3

```{r, collapse=TRUE}
# import dataset
ny_noaa_data = ny_noaa 
```

Data exploration 

```{r}

```

Short description, size and structure, key variables, indicating the extent to which missing data is an issue. 

### Questions

1. _*Do some data cleaning. Create separate variables for year, month, and day. Ensure observations for temperature, precipitation, and snowfall are given in reasonable units. For snowfall, what are the most commonly observed values? Why?*_

```{r}

```

2. _*Make a two-panel plot showing the average temperature in January and in July in each station across years. Is there any observable / interpretable structure? Any outliers?*_

```{r}

```

3. _*Make a two-panel plot showing (i) tmax vs tmin for the full dataset (note that a scatterplot may not be the best option); and (ii) make a plot showing the distribution of snowfall values greater than 0 and less than 100 separately by year.*_

```{r}

```





